---
title: "COVID-19 report"
author: "Daniel Levings"
date:  "`r format(Sys.time(), '%m/%d/%Y')`"
header-includes:
- \usepackage{pdflscape}
- \usepackage{pdfpages}
- \usepackage{float}
- \usepackage{xcolor}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[labelformat = empty]{caption}
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-14),Sys.Date(),'.html')) })
output:
    html_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.pos="H")
knitr::knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed = TRUE)})
knitr::opts_chunk$set(engine.path = list(python = '/home/daniel/anaconda3/envs/deeptools/bin/python'))
```  

## Purpose of this document/code:  

This is an **R Markdown** script created to document long-term trends in COVID-19 case and death rates across the United States and locally (in the Midwest/counties around Douglas Co, WI, where my family lives).  

```{r import packages, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library('doParallel')
library('BiocParallel')
library('splines')
library('magrittr')
library('tidyverse')
library('lubridate')
library('ggplot2')
library('plotly')
library('viridis')
library('readr')
library('tigris')
library('usmap')
library('rnaturalearth')
library('sf')
library('rgeos')
library('tidycensus')
library('extrafont')
loadfonts(device = 'postscript', quiet=TRUE)
options(tigris_use_cache = TRUE, Ncpus=floor(detectCores()*.9))
register(BPPARAM = MulticoreParam(floor(detectCores()*.9), progressbar = TRUE, exportglobals = FALSE))
options(MulticoreParam=MulticoreParam(workers = floor(detectCores()*0.9), 
    progressbar = TRUE, exportglobals = FALSE))
cores <- detectCores()

# If the script crashes, it's likely because the downloads take too long and timeout.  Try increasing
# size/number of seconds of the following variable
timeout_sec <- 800
```  

&nbsp;  

```{r get COVID and census data, results='hide'}
# Set API key for census data
# census_api_key("##############", install = T)

# Get variables that can be downloaded from the ACS dataset
ACS_vars <- load_variables(year = 2019, dataset = "acs1", cache = TRUE)
ACS_vars %>% filter(concept == "TOTAL POPULATION")

# Get population by county
Pop_size_by_cty <- get_estimates(geography = "county", product = "population", year = 2019) %>% 
    filter(variable %in% c("POP", "DENSITY")) %>% pivot_wider(id_cols = c("NAME", "GEOID"), 
    names_from = "variable", values_from = "value")

# Get data for United States counties
timeout_sec <- getOption('timeout')
if (file.exists("~/Desktop/COVID/COVID-19_NYT_data.csv")){
    NYT_data_date <- as.Date(file.info("~/Desktop/COVID/COVID-19_NYT_data.csv")$mtime)
} else {
    NYT_data_date <- as.Date(Sys.Date()-14)
}
if (as.numeric(Sys.Date() - NYT_data_date) > 1) {
    options(timeout = timeout_sec)
    download.file(url = "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv",
                  destfile = "~/Desktop/COVID/COVID-19_NYT_data.csv", method = "wget", quiet = T)
    options(timeout = timeout_sec)
} else {}
NYT.COVID19.data <- read_csv(file = "~/Desktop/COVID/COVID-19_NYT_data.csv", col_names = TRUE, col_types = cols(.default = col_character()))

# Get data from COVID19 John's Hopkins GitHub URL
if (file.exists("~/Desktop/COVID/COVID-19_JH_data.csv")){
    JH_data_date <- as.Date(file.info("~/Desktop/COVID/COVID-19_JH_data.csv")$mtime)
} else {
    JH_data_date <- as.Date(Sys.Date()-14)
}
if (as.numeric(Sys.Date() - JH_data_date) > 1) {
    options(timeout = timeout_sec)
    download.file(url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv",
                  destfile = "~/Desktop/COVID/COVID-19_JH_data.csv", method = "wget", quiet = T)
    options(timeout = timeout_sec)
} else {}
JH.COVID19.data <- read_csv(file = "~/Desktop/COVID/COVID-19_JH_data.csv", col_names = TRUE, col_types = cols(.default = col_character()))

# Convert numeric columns to numeric
NYT.COVID19.data[,5:ncol(NYT.COVID19.data)] <- sapply(NYT.COVID19.data[,5:ncol(NYT.COVID19.data)], as.numeric)
JH.COVID19.data[,3:ncol(JH.COVID19.data)] <- sapply(JH.COVID19.data[,3:ncol(JH.COVID19.data)], as.numeric)

# Extract data for countries/states of interest
USA.data <- NYT.COVID19.data %>% as.data.frame()
US.states <- USA.data[,"fips"] %>% unlist %>% as.character %>% unique %>% na.omit()
USA.data <- left_join(x=data.frame(fips = as.character(US.states)), y=USA.data[,c(1,4:ncol(USA.data))] %>%
    mutate(fips = as.character(fips)))

# All country data (in the process, clean up some countries' names)
Countries.data <- JH.COVID19.data %>% as.data.frame()
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Congo (Brazzaville)", "Congo (Kinshasa)", "Republic of the Congo")] <- "Democratic Republic of the Congo"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("The Bahamas", "Bahamas, The")] <- "Bahamas"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Brunei")] <- "Brunei Darussalam"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Cote d'Ivoire")] <- "Côte d'Ivoire"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Czechia")] <- "Czech Republic"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Taiwan*")] <- "Taiwan"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Russia")] <- "Russian Federation"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Korea, South")] <- "Republic of Korea"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Eswatini")] <- "Swaziland"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Holy See")] <- "Vatican"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("North Macedonia")] <- "Macedonia"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("Reunion")] <- "French Polynesia"
Countries.data$`Country/Region`[Countries.data$`Country/Region` %in% c("US")] <- "United States"
Countries.data <- Countries.data[,c(2,5:ncol(JH.COVID19.data))] %>% 
    group_by(`Country/Region`) %>% summarize(across(.fns = sum)) %>% as.data.frame()
Countries.names <- Countries.data[,1] %>% unlist %>% as.character()
Countries.data %<>% set_rownames(Countries.names)
Countries.data$`Country/Region` <- NULL

# Calculate local (last two weeks) average daily case and death rate
USA.summ <- USA.data %>% mutate(date = as.Date(date)) %>% arrange(fips, date) %>%
    # mutate(daily_cases = cases-lag(cases), daily_deaths = deaths-lag(deaths)) %>% 
    filter(date >= as.Date(Sys.Date()-14)) %>% 
    group_by(fips) %>% summarize(
    mean_cases_per_day = mean(cases-lag(cases), na.rm=T),
    mean_deaths_per_day = mean(deaths-lag(deaths), na.rm=T)
)

# Now get vaccination data - takes a few min
if (file.exists("~/Desktop/COVID/COVID-19_Vaccinations_in_the_United_States_County.tsv")){
    COVID_data_date <- as.Date(file.info("~/Desktop/COVID/COVID-19_Vaccinations_in_the_United_States_County.tsv")$mtime)
} else {
    COVID_data_date <- as.Date(Sys.Date()-14)
}
if (as.numeric(Sys.Date() - COVID_data_date) > 1) {
    options(timeout = timeout_sec)
    download.file(url = "https://data.cdc.gov/api/views/8xkx-amqh/rows.tsv?accessType=DOWNLOAD&bom=true",
                  destfile = "~/Desktop/COVID/COVID-19_Vaccinations_in_the_United_States_County.tsv", method = "wget", quiet = T)
    options(timeout = timeout_sec)
} else {}
CDC_vacc_data <- read.table(header = T, sep = "\t", quote = "", dec = ".", fill = T,
    file = "~/Desktop/COVID/COVID-19_Vaccinations_in_the_United_States_County.tsv")

# Now combine case and death rates, population numbers, and CDC vaccination rate data by county
USA.summ <- left_join(x=USA.summ, y=Pop_size_by_cty, by = c("fips" = "GEOID")) %>% 
    left_join(y=CDC_vacc_data %>% mutate(Date = as.Date(Date, format = "%m/%d/%Y")) %>% 
    arrange(FIPS, Date) %>% group_by(FIPS) %>% summarize(
    Vacc_Complete_Pct = as.numeric(max(Series_Complete_Pop_Pct, na.rm=T))+1,
    Vacc_FirstDose_Pct = as.numeric(max(Administered_Dose1_Pop_Pct, na.rm=T))+1),
              by = c("fips" = "FIPS"))

# add new function for scaling
stdize = function(x, ...) {(x - min(x, ...)) / (max(x, ...) - min(x, ...))}

# Add a factor comparing case/death and vaccination rates
USA.summ %<>% mutate(mean_cases_by_pop = mean_cases_per_day/(POP/100000),
    mean_deaths_by_pop = mean_deaths_per_day/(POP/100000),
    cases_vs_vacc = stdize(mean_cases_by_pop/(Vacc_FirstDose_Pct/101), na.rm=T),
    deaths_vs_vacc = stdize(mean_deaths_by_pop/(Vacc_FirstDose_Pct/101), na.rm=T),
    cases_vs_fullvacc = stdize(mean_cases_by_pop/(Vacc_Complete_Pct/101), na.rm=T),
    deaths_vs_fullvacc = stdize(mean_deaths_by_pop/(Vacc_Complete_Pct/101), na.rm=T))

USA.summ.quants <- rbind.data.frame(USA.summ %>% summarize(name = "cases_by_pop",
    `0` = min(mean_cases_by_pop, na.rm=T),
    `0.1` = quantile(mean_cases_by_pop, probs = 0.001, na.rm=T),
    `1` = quantile(mean_cases_by_pop, probs = 0.01, na.rm=T),
    `90` = quantile(mean_cases_by_pop, probs = 0.9, na.rm=T),
    `99` = quantile(mean_cases_by_pop, probs = 0.99, na.rm=T),
    `100` = max(mean_cases_by_pop, na.rm=T)),
    USA.summ %>% summarize(name = "deaths_by_pop",
    `0` = min(mean_deaths_by_pop, na.rm=T),
    `0.1` = quantile(mean_deaths_by_pop, probs = 0.001, na.rm=T),
    `1` = quantile(mean_deaths_by_pop, probs = 0.01, na.rm=T),
    `90` = quantile(mean_deaths_by_pop, probs = 0.9, na.rm=T),
    `99` = quantile(mean_deaths_by_pop, probs = 0.99, na.rm=T),
    `100` = max(mean_deaths_by_pop, na.rm=T)),
    USA.summ %>% summarize(name = "Vacc_FirstDose_Pct",
    `0` = min(Vacc_FirstDose_Pct, na.rm=T),
    `0.1` = quantile(Vacc_FirstDose_Pct, probs = 0.001, na.rm=T),
    `1` = quantile(Vacc_FirstDose_Pct, probs = 0.01, na.rm=T),
    `90` = quantile(Vacc_FirstDose_Pct, probs = 0.9, na.rm=T),
    `99` = quantile(Vacc_FirstDose_Pct, probs = 0.99, na.rm=T),
    `100` = max(Vacc_FirstDose_Pct, na.rm=T)),
    USA.summ %>% summarize(name = "Vacc_Complete_Pct",
   `0` = min(Vacc_Complete_Pct, na.rm=T),
   `0.1` = quantile(Vacc_Complete_Pct, probs = 0.001, na.rm=T),
   `1` = quantile(Vacc_Complete_Pct, probs = 0.01, na.rm=T),
   `90` = quantile(Vacc_Complete_Pct, probs = 0.9, na.rm=T),
   `99` = quantile(Vacc_Complete_Pct, probs = 0.99, na.rm=T),
   `100` = max(Vacc_Complete_Pct, na.rm=T)),
    USA.summ %>% summarize(name = "cases_vs_vacc",
   `0` = min(cases_vs_vacc, na.rm=T),
   `0.1` = quantile(cases_vs_vacc, probs = 0.001, na.rm=T),
   `1` = quantile(cases_vs_vacc, probs = 0.01, na.rm=T),
   `90` = quantile(cases_vs_vacc, probs = 0.9, na.rm=T),
   `99` = quantile(cases_vs_vacc, probs = 0.99, na.rm=T),
   `100` = max(cases_vs_vacc, na.rm=T)),
    USA.summ %>% summarize(name = "deaths_vs_vacc",
    `0` = min(deaths_vs_vacc, na.rm=T),
    `0.1` = quantile(deaths_vs_vacc, probs = 0.001, na.rm=T),
    `1` = quantile(deaths_vs_vacc, probs = 0.01, na.rm=T),
    `90` = quantile(deaths_vs_vacc, probs = 0.9, na.rm=T),
    `99` = quantile(deaths_vs_vacc, probs = 0.99, na.rm=T),
    `100` = max(deaths_vs_vacc, na.rm=T)),
    USA.summ %>% summarize(name = "cases_vs_fullvacc",
    `0` = min(cases_vs_fullvacc, na.rm=T),
    `0.1` = quantile(cases_vs_fullvacc, probs = 0.001, na.rm=T),
    `1` = quantile(cases_vs_fullvacc, probs = 0.01, na.rm=T),
    `90` = quantile(cases_vs_fullvacc, probs = 0.90, na.rm=T),
    `99` = quantile(cases_vs_fullvacc, probs = 0.99, na.rm=T),
    `100` = max(cases_vs_fullvacc, na.rm=T)),
    USA.summ %>% summarize(name = "deaths_vs_fullvacc",
    `0` = min(deaths_vs_fullvacc, na.rm=T),
    `0.1` = quantile(deaths_vs_fullvacc, probs = 0.001, na.rm=T),
    `1` = quantile(deaths_vs_fullvacc, probs = 0.001, na.rm=T),
    `90` = quantile(deaths_vs_fullvacc, probs = 0.9, na.rm=T),
    `99` = quantile(deaths_vs_fullvacc, probs = 0.99, na.rm=T),
    `100` = max(deaths_vs_fullvacc, na.rm=T)))

# Now isolate/extract only data for WI and MN
# First reorganize
USA.summ <- USA.summ %>% mutate(State = trimws(gsub(x=NAME, pattern = "^.*, ?", replacement = "", perl = T))) %>%
    dplyr::select(1, NAME, State, POP, DENSITY, 2:3, 7:length(colnames(USA.summ)))

# Then get appropriate data
MN_WI_data <- USA.summ %>% filter(State %in% c("Wisconsin", "Minnesota"))

# Get mins and maxes
MN_WI_data.quants <- rbind.data.frame(USA.summ %>% summarize(name = "cases_by_pop",
      `0` = min(mean_cases_by_pop, na.rm=T),
      `0.1` = quantile(mean_cases_by_pop, probs = 0.001, na.rm=T),
      `1` = quantile(mean_cases_by_pop, probs = 0.01, na.rm=T),
      `90` = quantile(mean_cases_by_pop, probs = 0.9, na.rm=T),
      `95` = quantile(mean_cases_by_pop, probs = 0.95, na.rm=T),
      `99` = quantile(mean_cases_by_pop, probs = 0.99, na.rm=T),
      `100` = max(mean_cases_by_pop, na.rm=T)),
    MN_WI_data %>% summarize(name = "deaths_by_pop",
      `0` = min(mean_deaths_by_pop, na.rm=T),
      `0.1` = quantile(mean_deaths_by_pop, probs = 0.001, na.rm=T),
      `1` = quantile(mean_deaths_by_pop, probs = 0.01, na.rm=T),
      `90` = quantile(mean_deaths_by_pop, probs = 0.9, na.rm=T),
      `95` = quantile(mean_deaths_by_pop, probs = 0.95, na.rm=T),
      `99` = quantile(mean_deaths_by_pop, probs = 0.99, na.rm=T),
      `100` = max(mean_deaths_by_pop, na.rm=T)),
    MN_WI_data %>% summarize(name = "Vacc_FirstDose_Pct",
      `0` = min(Vacc_FirstDose_Pct, na.rm=T),
      `0.1` = quantile(Vacc_FirstDose_Pct, probs = 0.001, na.rm=T),
      `1` = quantile(Vacc_FirstDose_Pct, probs = 0.01, na.rm=T),
      `90` = quantile(Vacc_FirstDose_Pct, probs = 0.9, na.rm=T),
      `95` = quantile(Vacc_FirstDose_Pct, probs = 0.95, na.rm=T),
      `99` = quantile(Vacc_FirstDose_Pct, probs = 0.99, na.rm=T),
      `100` = max(Vacc_FirstDose_Pct, na.rm=T)),
    MN_WI_data %>% summarize(name = "Vacc_Complete_Pct",
      `0` = min(Vacc_Complete_Pct, na.rm=T),
      `0.1` = quantile(Vacc_Complete_Pct, probs = 0.001, na.rm=T),
      `1` = quantile(Vacc_Complete_Pct, probs = 0.01, na.rm=T),
      `90` = quantile(Vacc_Complete_Pct, probs = 0.9, na.rm=T),
      `95` = quantile(Vacc_Complete_Pct, probs = 0.95, na.rm=T),
      `99` = quantile(Vacc_Complete_Pct, probs = 0.99, na.rm=T),
      `100` = max(Vacc_Complete_Pct, na.rm=T)),
    MN_WI_data %>% summarize(name = "cases_vs_vacc",
      `0` = min(cases_vs_vacc, na.rm=T),
      `0.1` = quantile(cases_vs_vacc, probs = 0.001, na.rm=T),
      `1` = quantile(cases_vs_vacc, probs = 0.01, na.rm=T),
      `90` = quantile(cases_vs_vacc, probs = 0.9, na.rm=T),
      `95` = quantile(cases_vs_vacc, probs = 0.95, na.rm=T),
      `99` = quantile(cases_vs_vacc, probs = 0.99, na.rm=T),
      `100` = max(cases_vs_vacc, na.rm=T)),
    MN_WI_data %>% summarize(name = "deaths_vs_vacc",
      `0` = min(deaths_vs_vacc, na.rm=T),
      `0.1` = quantile(deaths_vs_vacc, probs = 0.001, na.rm=T),
      `1` = quantile(deaths_vs_vacc, probs = 0.01, na.rm=T),
      `90` = quantile(deaths_vs_vacc, probs = 0.9, na.rm=T),
      `95` = quantile(deaths_vs_vacc, probs = 0.95, na.rm=T),
      `99` = quantile(deaths_vs_vacc, probs = 0.99, na.rm=T),
      `100` = max(deaths_vs_vacc, na.rm=T)),
    MN_WI_data %>% summarize(name = "cases_vs_fullvacc",
      `0` = min(cases_vs_fullvacc, na.rm=T),
      `0.1` = quantile(cases_vs_fullvacc, probs = 0.001, na.rm=T),
      `1` = quantile(cases_vs_fullvacc, probs = 0.01, na.rm=T),
      `90` = quantile(cases_vs_fullvacc, probs = 0.90, na.rm=T),
      `95` = quantile(cases_vs_fullvacc, probs = 0.95, na.rm=T),
      `99` = quantile(cases_vs_fullvacc, probs = 0.99, na.rm=T),
      `100` = max(cases_vs_fullvacc, na.rm=T)),
    MN_WI_data %>% summarize(name = "deaths_vs_fullvacc",
      `0` = min(deaths_vs_fullvacc, na.rm=T),
      `0.1` = quantile(deaths_vs_fullvacc, probs = 0.001, na.rm=T),
      `1` = quantile(deaths_vs_fullvacc, probs = 0.001, na.rm=T),
      `90` = quantile(deaths_vs_fullvacc, probs = 0.9, na.rm=T),
      `95` = quantile(deaths_vs_fullvacc, probs = 0.95, na.rm=T),
      `99` = quantile(deaths_vs_fullvacc, probs = 0.99, na.rm=T),
      `100` = max(deaths_vs_fullvacc, na.rm=T))) 
```  

&nbsp;  
&nbsp;  

Plot of country-wide data/trends for USA:  

```{r plot USA, fig.width=13, fig.height=8}
plot_usmap(data = USA.summ, values = "mean_deaths_by_pop", color = "black") + 
    scale_fill_viridis(option="A", name = "Mean deaths per\nday per 100K", direction=-1, oob = scales::squish, 
      limits = c(USA.summ.quants %>% filter(name == "deaths_by_pop") %>% dplyr::select(`1`) %>% unlist(),
      USA.summ.quants %>% filter(name == "deaths_by_pop") %>% dplyr::select(`99`) %>% unlist())) + 
    ggtitle(paste("2021 COVID-19 two-week average death rate - as of",
                  Sys.Date())) +
    theme(legend.position = "right", 
          plot.title = element_text(face = "bold", size = 14))
plot_usmap(data = USA.summ, values = "mean_cases_by_pop", color = "black") + 
    scale_fill_viridis(option="A", name = "Mean cases per\nday per 100K", direction=-1, oob = scales::squish, 
                       limits = c(USA.summ.quants %>% filter(name == "cases_by_pop") %>% dplyr::select(`1`) %>% unlist(),
                                  USA.summ.quants %>% filter(name == "cases_by_pop") %>% dplyr::select(`99`) %>% unlist())) + 
    ggtitle(paste("2021 COVID-19 two-week average case rate - as of",
                  Sys.Date())) +
    theme(legend.position = "right", 
          plot.title = element_text(face = "bold", size = 14))
plot_usmap(data = USA.summ, values = "Vacc_FirstDose_Pct", color = "black") + 
    scale_fill_viridis(option="D", name = "First Dose Pct%", trans = "log2", direction=-1,
                       breaks = c(2,5,17,64), labels = c(1,4,16,64)) + 
    ggtitle(paste("2021 COVID-19 vaccination percent - as of",
                  Sys.Date())) +
    theme(legend.position = "right", 
          plot.title = element_text(face = "bold", size = 14))
```  

&nbsp;  
&nbsp;  

Now plot of statewide (MN and WI) data/trends:  

```{r plot statewide, fig.width=13, fig.height=8}
plot_usmap(data = MN_WI_data, values = "mean_deaths_by_pop", color = "black", 
     include = MN_WI_data %>% dplyr::select(fips) %>% unlist %>% unique()) + 
  scale_fill_viridis(option="A", name = "Mean deaths per\nday per 100K", direction=-1, oob = scales::squish, 
     limits = c(MN_WI_data.quants %>% filter(name == "deaths_by_pop") %>% dplyr::select(`1`) %>% unlist(),
      MN_WI_data.quants %>% filter(name == "deaths_by_pop") %>% dplyr::select(`95`) %>% unlist())) + 
  ggtitle(paste("2021 COVID-19 two-week average death rate - as of",
                Sys.Date())) +
  theme(legend.position = "right", 
      plot.title = element_text(face = "bold", size = 14))
plot_usmap(data = MN_WI_data, values = "mean_cases_by_pop", color = "black", 
       include = MN_WI_data %>% dplyr::select(fips) %>% unlist %>% unique()) + 
  scale_fill_viridis(option="A", name = "Mean cases per\nday per 100K", direction=-1, oob = scales::squish, 
   limits = c(MN_WI_data.quants %>% filter(name == "cases_by_pop") %>% dplyr::select(`1`) %>% unlist(),
    MN_WI_data.quants %>% filter(name == "cases_by_pop") %>% dplyr::select(`95`) %>% unlist())) + 
  ggtitle(paste("2021 COVID-19 two-week average case rate - as of",
                Sys.Date())) +
  theme(legend.position = "right", 
    plot.title = element_text(face = "bold", size = 14))
plot_usmap(data = MN_WI_data, values = "Vacc_FirstDose_Pct", color = "black", 
    include = MN_WI_data %>% dplyr::select(fips) %>% unlist %>% unique()) + 
  scale_fill_viridis(option="D", name = "First Dose Pct%", direction=-1) + 
  ggtitle(paste("2021 COVID-19 vaccination percent - as of",
                Sys.Date())) +
  theme(legend.position = "right", 
    plot.title = element_text(face = "bold", size = 14))
```  

&nbsp;  
&nbsp;  

Plot of local trends in case rate:  

```{r plot local trends}
# Now get the data for the case rate in all counties nearby Douglas
# Change to your state/nearby states and counties here:
Local_COVID_data <- NYT.COVID19.data %>% 
    filter(((county %in% c("Pine", "Carlton", "St. Louis")) & (state %in% "Minnesota")) | 
    ((county %in% c("Douglas", "Bayfield", "Sawyer", "Washburn", "Ashland", "Burnett")) & 
    (state %in% "Wisconsin"))) %>% mutate(date = as.Date(date)) %>% arrange(fips, date) %>%
    mutate(cases_per_day = cases-lag(cases)) %>% filter(date >= as.Date(Sys.Date()-(16*7))) 
Local_COVID_data <- Local_COVID_data %>% left_join(y=Pop_size_by_cty, by = c("fips" = "GEOID")) %>%
    mutate(new_cases_per_100k = cases_per_day/(POP/100000))

# Remove weekends (always 0's)
Local_COVID_data <- Local_COVID_data %>% mutate(weekday = wday(date)) %>% filter((weekday > 1) & (weekday < 7))

# Get plot boundaries
plot_max <- ceiling(as.numeric(quantile(Local_COVID_data$cases_per_day, probs = c(0.995)))/10)*10
plot_max2 <- ceiling(as.numeric(quantile(Local_COVID_data$new_cases_per_100k, probs = c(0.99)))/10)*10 

# Now plot the trends of new cases for the last ~3 months
p1 <- Local_COVID_data %>% 
    ggplot(mapping = aes(x = date, y = cases_per_day, group = county, fill = county, color = county)) +
    geom_point(alpha = 0.35) + stat_smooth(geom="smooth", method = "lm", formula = y ~ bs(x=x, df=3), se = F, linetype = 2) + 
    scale_fill_manual(values = rainbow(9)[c(rbind(seq(1,7,2), c(seq(6,9,2),seq(2,5,2))), 9)]) +
    scale_color_manual(values = rainbow(9)[c(rbind(seq(1,7,2), c(seq(6,9,2),seq(2,5,2))), 9)]) +
    ggtitle(paste0("Nearby COVID-19 cases as of ", as.character(Sys.Date()))) + 
    coord_cartesian(xlim = c(as.Date(Sys.Date()-(12*7)), as.Date(Sys.Date())), ylim = c(0, plot_max), expand = F) +
    # scale_y_log10() + 
    theme_bw() +
    theme(plot.title = element_text(face="bold", hjust=0.5, size=16), 
          axis.text = element_text(color="black", size=12),
          axis.title = element_text(face="bold", color="black", size=14), 
          axis.ticks = element_line(color="black"),
          panel.border = element_rect(size=0.5, color="black"),
          legend.text = element_text(color="black", size=12), 
          legend.title = element_text(color = "black", face = "bold.italic", size = 13),
          plot.margin = unit(c(5.5,9,5.5,5.5), "points"),
          panel.grid = element_line(color = "gray95", linetype = 2)
    )
p2 <- Local_COVID_data %>% 
    ggplot(mapping = aes(x = date, y = new_cases_per_100k, group = county, fill = county, color = county)) +
    geom_point(alpha = 0.35) + stat_smooth(geom="smooth", method = "lm", formula = y ~ bs(x=x, df=3), se = F, linetype = 2) + 
    scale_fill_manual(values = rainbow(9)[c(rbind(seq(1,7,2), c(seq(6,9,2),seq(2,5,2))), 9)]) +
    scale_color_manual(values = rainbow(9)[c(rbind(seq(1,7,2), c(seq(6,9,2),seq(2,5,2))), 9)]) +
    ggtitle(paste0("Nearby COVID-19 cases as of ", as.character(Sys.Date()))) + 
    coord_cartesian(xlim = c(as.Date(Sys.Date()-(12*7)), as.Date(Sys.Date())), ylim = c(0, plot_max2), expand = F) +
    # scale_y_log10() + 
    theme_bw() +
    theme(plot.title = element_text(face="bold", hjust=0.5, size=16), 
        axis.text = element_text(color="black", size=12),
        axis.title = element_text(face="bold", color="black", size=14), 
        axis.ticks = element_line(color="black"),
        panel.border = element_rect(size=0.5, color="black"),
        legend.text = element_text(color="black", size=12), 
        legend.title = element_text(color = "black", face = "bold.italic", size = 13),
        plot.margin = unit(c(5.5,9,5.5,5.5), "points"),
        panel.grid = element_line(color = "gray95", linetype = 2)
    )
```  

```{r plotly, fig.width=13, fig.height=8}
m <- list(l = 50, r = 50, b = 100, t = 100, pad = 4)
p1 <- ggplotly(p1, width = 1000, height = 700) %>% layout(autosize = F, margin = m)
p2 <- ggplotly(p2, width = 1000, height = 700) %>% layout(autosize = F, margin = m)
p1
```  

&nbsp;  

```{r plotly 2, fig.width=13, fig.height=8}
p2
```  
